### 2014-03-20 manuscript.tex
Described dependency distance, added paragraph headings to section 2.2

\subsection{Parse tree method}
\label{sec:dp}
%LILJA: OM PARSING
When determining the sentiment expressed towards a specific target word, the syntactic environment of this word and how it relates to sentiment-bearing words in the context may clearly be of importance.

\paragraph{Dependency distance}
One way of expressing the syntactic environment of a target word with respect to a sentiment word is to determine its distance in the dependency graph. We define a distance function $\mathbf{depdist}(kw_i, sw_{j})$ which returns the number of nodes between the two words in the parse tree plus one.

LILJA:OM DEPENDENCY DISTANCE?

A second way of determining the importance of a sentiment word towards a target based on syntactically parsed texts, is to establish a list of grammatical dependency paths between words, and test whether such paths exist between the target and the sentiment words in the material under investigation \cite{Jiang11}. The assumption would be that, if there is a meaningful grammatical relation between a target and a sentiment word, it is likely that they are semantically related to each other. Furthermore, it is reasonable to expect than some paths are stronger indicators of the overall sentiment of the sentence than others. To test this method, we have made a list of 42 grammatical dependency paths and given them a score between 1-3. The higher the score is, the better indicator of sentiment the path is assumed to be. In the following paragraphs, we will present our reasoning behind the choice of paths and the weight we have given them. The paths are written as follows: postag-target:postag-sentiment word{\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn({\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn etc.). \emph{Up} and \emph{dn} show whether you move up or down in the dependency tree. So the relation between a noun target and a verb sentiment word where the noun is the subject of the verb, is represented as subst:verb{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up. The relation between a target noun and a sentiment noun where the former is the subject and the latter is the direct object of the same verb, is subst:subst{\textunderscore}{\textunderscore}SUBST{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn.

@@

Let $\mathcal{D}$ denote the set of all important grammatical relations. The function $\mathbf{gram}(kw_i, sw_{j})$ returns the grammatical relation, and if $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$, then the function $W_{\mathbf{dep}}(kw_i, sw_{j}) \in [0,1]$, returns the importance of the grammatical relation. Further let $\mathbf{treedist}(kw_i, sw_{j})$ return the number of words between the two words in the parse tree plus one. The \textbf{imp} function is computed as follows. If $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$ we use
\begin{align}
  \begin{split}
    \label{eq:3}
  \mathbf{imp}&(kw_i, sw_{j}) = \\
  &\alpha \cdot \mathbf{sentsc}(sw_{j}) W_{\mathbf{dep}}(kw_i, sw_{j}) \\
  +&(1 - \alpha) \cdot \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
  \end{split}
\end{align}
where $\alpha \in [0,1]$ is a parameter that weights the score from the important dependence path and the tree distance and can be tuned using a training set. If $\mathbf{gram}(kw_i, sw_{j}) \not\in \mathcal{D}$ we simply use 
\begin{equation}
  \label{eq:4}
  \mathbf{imp}(kw_i, sw_{j}) = \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
\end{equation}
Note that when $\alpha = 0$, \eqref{eq:3} reduces to \eqref{eq:4}.

>>>>

\subsection{Parse tree method}
\label{sec:dp}
%LILJA: OM PARSING
When determining the sentiment expressed towards a specific target
word, the syntactic environment of this word and how it relates to
sentiment-bearing words in the context may clearly be of importance.
In the following we present a modification of the scoring function described above to also take into account the syntactic environment of the target words. The function is defined over dependency graphs, i.e. connected, acyclic graphs expressing bilexical relations.

\paragraph{Dependency distance} One way of expressing the syntactic environment of a target word with respect to a sentiment word is to determine its distance in the dependency graph. We therefore define a distance function $\mathbf{depdist}(kw_i, sw_{j})$ which returns the number of nodes in the shortest dependency path from the target word to the sentiment word in the dependency.

\paragraph{Dependency paths}
A second way of determining the importance of a sentiment word towards a target based on syntactically parsed texts, is to establish a list of grammatical dependency paths between words, and test whether such paths exist between the target and the sentiment words in the material under investigation \cite{Jiang11}. The assumption would be that, if there is a meaningful grammatical relation between a target and a sentiment word, it is likely that they are semantically related to each other. Furthermore, it is reasonable to expect than some paths are stronger indicators of the overall sentiment of the sentence than others. To test this method, we have made a list of 42 grammatical dependency paths and given them a score between 1-3. The higher the score is, the better indicator of sentiment the path is assumed to be. In the following paragraphs, we will present our reasoning behind the choice of paths and the weight we have given them. The paths are written as follows: postag-target:postag-sentiment word{\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn({\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn etc.). \emph{Up} and \emph{dn} show whether you move up or down in the dependency tree. So the relation between a noun target and a verb sentiment word where the noun is the subject of the verb, is represented as subst:verb{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up. The relation between a target noun and a sentiment noun where the former is the subject and the latter is the direct object of the same verb, is subst:subst{\textunderscore}{\textunderscore}SUBST{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn.

@@

\paragraph{Modified scoring function}
Let $\mathcal{D}$ denote the set of all important grammatical relations. The function $\mathbf{gram}(kw_i, sw_{j})$ returns the grammatical relation, and if $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$, then the function $W_{\mathbf{dep}}(kw_i, sw_{j}) \in [0,1]$, returns the importance of the grammatical relation. Further let $\mathbf{depdist}(kw_i, sw_{j})$ return the dependency distance, as described above. The \textbf{imp} function is computed as follows. If $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$ we use
\begin{align}
  \begin{split}
    \label{eq:3}
  \mathbf{imp}&(kw_i, sw_{j}) = \\
  &\alpha \cdot \mathbf{sentsc}(sw_{j}) W_{\mathbf{dep}}(kw_i, sw_{j}) \\
  +&(1 - \alpha) \cdot \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
  \end{split}
\end{align}
where $\alpha \in [0,1]$ is a parameter that weights the score from the important dependence path and the tree distance and can be tuned using a training set. If $\mathbf{gram}(kw_i, sw_{j}) \not\in \mathcal{D}$ we simply use 
\begin{equation}
  \label{eq:4}
  \mathbf{imp}(kw_i, sw_{j}) = \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
\end{equation}
Note that when $\alpha = 0$, \eqref{eq:3} reduces to \eqref{eq:4}.
